Pure Spin Current Injection in Hydrogenated Graphene Structures

### 2017-07-14 mu/article/article-review4_1.tex
Adding new plots to better explain the results.

\documentclass[prb,11pt,tightenlines,twocolumn,aps]{revtex4-1}

\usepackage{amsmath}    % need for subequations
\usepackage{graphicx}   % need for figures
\usepackage{verbatim}   % useful for program listings
\usepackage{color}      % use if color is used in text
\usepackage{subfigure}  % use for side-by-side figures
\usepackage{hyperref}   % use for hypertext links, including those to external
                        % documents and URLs 
\usepackage{blindtext}  % fill text
\usepackage[dvipsnames]{xcolor}
\usepackage{multirow}
\usepackage[capitalize]{cleveref}
% \usepackage{showframe}
% \usepackage[inline]{showlabels}

@@

\begin{abstract}
\blindtext
\end{abstract}

@@

In PSCs there is no net motion of electrical charge and spin-up electrons move
in a given direction while spin-down electrons travel in the opposite one.
This effect can result from one photon absorption of linearly polarized light
by a semiconductor, with filled valence bands and empty conduction bands,
illuminated by light with photon energy larger than the energy gap.
% note 1
Using a single continuous linearly polarized laser beam, it is possible
to promote electrons in $\mathbf{k}$ space regardless the symmetry of the
system resulting in a net current equal to zero. 
% 
If the phenomena is produced in a noncentrosymmetric semiconducting media, the
electrons promoted to the conduction bands at opposite $\mathbf{k}$ points
can produce a net spin polarization different to zero\cite{alvaradoPRL85},
resulting in a PSC because the velocity of electrons
at opposite $\mathbf{k}$ points are in opposite directions.

Using the multiple scale approach to solve the equation of motion for the
single particle density matrix $\rho_{mn}(\mathbf{k};t)$ we have that 
\begin{equation*}
\begin{aligned}
&\frac{\partial \rho_{cc'}}{\partial t} =
-i (\omega_{cc'} - i \epsilon) \rho_{cc'}
+
\frac{e^{2}E^{\mathrm{a}}(\omega)E^{\mathrm{b*}}(\omega)}
{i \hbar^{2}}
\times \\
& \quad \sum_{v}r^{\mathrm{a}}_{cv} r^{\mathrm{b}}_{vc'}
\left( \frac{1}{\omega - \omega_{c'v} - i \epsilon} 
- 
\frac{1}{\omega - \omega_{cv} + i \epsilon} \right).
\end{aligned}
\end{equation*}
Taking in the first term $\epsilon \rightarrow 0$ and changing the density
matrix operator to the interaction representation, $\tilde{\hat{\rho}} = e^{iH_
{0}t/\hbar} \hat{\rho} e^{-iH_{0}t/\hbar}$, with $H_{0}$ the ground state
Hamiltonian, the matrix elements can be written as
\begin{equation*}
\tilde{\rho}_{cc'(\mathbf{k})} 
= \langle c\mathbf{k}| \tilde{\hat{\rho}} | c'\mathbf{k} \rangle
= e^{i \omega_{cc'}t} \rho_{cc'}(\mathbf{k}).
\end{equation*}
where $H_{0}|n\mathbf{k}\rangle = \hbar \omega_{n}(\mathbf{k})|n
\mathbf{k}\rangle$ with $\hbar \omega_{n}(\mathbf{k})$ the energy of the
electronic band $n$ at point $\mathbf{k}$ in the irreducible Brillouin zone
(IBZ), and $|n\mathbf{k}\rangle$ the Bloch state. Then, the time derivative of
$\tilde{\rho}_{cc'}(\mathbf{k})$ is given by 
\begin{equation}\label{eq:derden}
\begin{aligned}
&\frac{d\tilde{\rho}_{cc'}(\mathbf{k})}{dt} 
= 
\frac{e^{2} E^\mathrm{a}(\omega) E^{\mathrm{b*}}(\omega)}{i\hbar^{2}} 
e^{i\omega_{cc'}t} 
\times \\
&\quad \sum_{v}r^{\mathrm{a}}_{cv} r^{\mathrm{b}}_{vc'}
\left( \frac{1}{\omega - \omega_{c'v} - i \epsilon} 
- 
\frac{1}{\omega - \omega_{cv} + i \epsilon} \right),
\end{aligned}
\end{equation}
where $\epsilon \rightarrow 0$ still needs to be taken. 
% 
The expectation value of an observable $\mathcal{O}$ with quantum mechanical
operator associated $\hat{\mathcal{O}}$ is given by
\begin{equation}\label{eq:trace}
\mathcal{O} = \mathrm{Tr}(\hat{\rho}\hat{\mathcal{O}}),
\end{equation}
where $\mathrm{Tr}$ denotes the trace. Then
\begin{align}
\mathcal{O} 
=&
\int \frac{d^{3}k}{8\pi^{3}}\sum_{c} \langle c \mathbf{k} 
| \hat{\rho} \hat{\mathcal{O}} |
c\mathbf{k} \rangle \nonumber \\
=& 
\int \frac{d^{3}k}{8\pi^{3}} \sum_{cc'} \tilde{\rho}_{cc'}(\mathbf{k}) 
\mathcal{O}_{c'c}(\mathbf{k}),
\label{eq:expectation}
\end{align}
where we used the interaction picture representation,
$\tilde{\mathcal{O}}_{cc'} = \langle c'\mathbf{k} | e^{iH_{0}t/\hbar}
\hat{\mathcal{O}} e^{-iH_{0}t/\hbar} | c\mathbf{k} \rangle$, and the closure
relationship $\sum_{c}|c\mathbf{k}\rangle \langle c\mathbf{k} = 1$. So, the
rate of change of $\mathcal{O}$ can be written from Eq. \eqref{eq:derden}, Eq.
\eqref{eq:trace}, and Eq. \eqref{eq:expectation} as
\begin{align}
\dot{\mathcal{O}} 
&=
\mathrm{Tr} \left( \frac{d \hat{\tilde{\rho}}}{dt} 
\tilde{\mathcal{O}} \right), 
\nonumber \\
&=\frac{e^{2}}{i\hbar^{2}} \int \frac{d^{3}k}{8\pi^{3}} 
\sum \mathcal{O}_{c'c} r^{\mathrm{a}}_{cv} r^{\mathrm{b}}_{vc'} \times
\nonumber \\
& \left( \frac{1}{\omega - \omega_{c'v} - i\epsilon} - 
\frac{1}{\omega - \omega_{cv} + i\epsilon} \right)
E^{\mathrm{a}}(\omega) E^{\mathrm{b*}}(\omega)
\label{eq:dotO}
\end{align}

Replacing the operator $\mathcal{O} \rightarrow \hat{K}^{\mathrm{ab}} \equiv v^
{\mathrm{a}} S^{\mathrm{b}}$ the matrix elements are given by
\begin{equation}
K^{\mathrm{ab}}_{nm}(\mathbf{k}) = 
\sum_{l} v^{\mathrm{a}}_{nl}(\mathbf{k})S^{\mathrm{b}}_{lm}(\mathbf{k}), 
\label{eq:velspimatelem}
\end{equation}
and using time-reversal invariance it follows that 
\begin{equation}
K^{\mathrm{ab}}_{nm}(\mathbf{-k}) = K^{\mathrm{ab*}}_{nm}(\mathbf{k}). 
\label{eq:k-timerev}
\end{equation}
Now the time derivative $\dot{K}^\mathrm{ab}$ can be obtained as we proceeded
in Eq. \eqref{eq:dotk-dev} of Appendix \ref{sec:mathder} and then we write
\begin{equation}
\dot{K}^{\mathrm{ab}}(\omega) =
\mu^{\mathrm{abcd}}(\omega)
E^{\mathrm{c}}(\omega) E^{\mathrm{d*}}(\omega),
\label{eq:dotk}
\end{equation}
where the pure spin-current pseudotensor components
\begin{equation}\label{eq:mu}
\begin{aligned}
\mu^{\mathrm{abcd}}  (\omega) &
=
\frac{\pi e^{2}}{\hbar^{2}} \int 
\frac{d^{3}K}{8 \pi^{3}} \times \\
\sum_{vcc'}^{'}
\mathrm{Re} & \left[ K^{\mathrm{ab}}_{cc'} 
\left( 
r^{\mathrm{c}}_{vc'} 
r^{\mathrm{d}}_{cv } +
(c \leftrightarrow d)
\right) 
\right] \delta(\omega-\omega_{cv}),
\end{aligned}
\end{equation}
are obtained from Eq. \eqref{eq:k-final} of the Appendix.
% 
The $'$ in the sum means that $c$ and $c'$ are quasi degenerate states and the
sum only covers these states and since $\mu^{\mathrm{abcd}}(\omega)$ is real we
have that $\mu^{\mathrm{abcd}}(\omega) = \mu^{\mathrm{abdc}}(\omega)$. This
previous equation is the same as Eq. (3) of Bhat et al.\cite{bhatPRL05}
obtained by using the semiconductor optical Bloch equations.

@@

\subsection{Spin velocity injection} % (fold)
\label{sec:theory-pure_spin_current}

We define the SVI as the velocity at which the spin, polarized along the
direction $\mathrm{a}$, propagates along the direction $\mathrm{b}$ as
\begin{equation}\label{eq:vab-w}
\mathcal{V}^{\mathrm{ab}}(\omega) \equiv
\frac{\dot{K}^{\mathrm{ab}}(\omega)}{(\hbar/2) \dot{n}(\omega)},
\end{equation}  
where $\dot{K}^{\mathrm{ab}} (\omega)$ is given by Eqns. \eqref{eq:dotk} and
\eqref{eq:mu} and the carrier
injection rate,
$\dot{n}(\omega)$, are given by
% 
\begin{equation}
\dot{n}(\omega) =
\xi^{\mathrm{ab}}(\omega) E^{c }(\omega) E^{d*}(\omega),
\label{eq:dotn}
\end{equation}
where $\xi^{\mathrm{ab}}(\omega)$ are the carrier
generation rate tensor components. 
% 
Since we have 2D structures we use an incoming electric field parallel to the
surface, $\mathbf{E}^{\mathrm{a}}(\omega,\alpha) = E^{\mathrm{a}}(\omega)
e^{i\alpha}$, where $\alpha$ corresponds to the linear polarization angle.
% 
Then from Eq. \eqref{eq:dotk}, Eq. \eqref{eq:vab-w}, and \eqref{eq:dotn}, and
including the polarization angle dependence we obtain
\begin{widetext}
\begin{align}
\mathcal{V}^{\mathrm{ab}}(\omega,\alpha)
% &= 
% \frac{2}{\hbar}
% \frac{\mu^{\mathrm{abxx}}(\omega)
% E^{2}(\omega)\cos^{2}(\alpha) + 
% \mu^{\mathrm{abyy}}(\omega)
% E^{2}(\omega)\sin^{2}(\alpha) + 
% 2\mu^{\mathrm{abxy}}(\omega)
% E^{2}(\omega)\cos(\alpha)\sin(\alpha)}
% {\xi^{\mathrm{xx}}(\omega)
% E^{2}(\omega)\cos^{2}(\alpha) + 
% \xi^{\mathrm{yy}}(\omega)
% E^{2}(\omega)\sin^{2}(\alpha)},
% \nonumber \\
&= 
\frac{2}{\hbar}
\frac{\mu^{\mathrm{abxx}}(\omega)\cos^{2}(\alpha) + 
\mu^{\mathrm{abyy}}(\omega)\sin^{2}(\alpha) + 
\mu^{\mathrm{abxy}}(\omega)\sin(2\alpha)}
{\xi^{\mathrm{xx}}(\omega)\cos^{2}(\alpha) + 
\xi^{\mathrm{yy}}(\omega)\sin^{2}(\alpha)},
\label{eq:vab-aw}
\end{align}
\end{widetext}
% 
Two interesting possibilities to analyze the SVI are fixing the spin along $z$,
directed perpendicularly to the surface of the structure, or fixing the
velocity along $x$ or $y$ on the $xy$ plane of the structures. Also we can
analyze the SVI contribution coming from each layer of the structure. In
following subsections we present these three cases.

@@

Analyzing the SVI, Eq. \eqref{eq:vab-aw}, we define the magnitude of the spin
velocity with spin polarized along the $\mathrm{b}$ direction as
\begin{equation}
|\mathcal{V}_{\sigma^{\mathrm{b}}}(\omega,\alpha)| 
=
\sqrt{
[\mathcal{V}^{\mathrm{xb}}(\omega,\alpha)]^{2}\ +
[\mathcal{V}^{\mathrm{yb}}(\omega,\alpha)]^{2}\ 
}, 
\label{eq:vs-mag}
\end{equation}
and the angle at which the spin velocity is directed on the $xy$ plane as
\begin{equation}
\gamma_{\mathrm{b}} (\omega,\alpha)
=
\tan^{-1} \left( \frac{\mathcal{V}^{\mathrm{yb}}(\omega,\alpha)}
{\mathcal{V}^{\mathrm{xb}}(\omega,\alpha)} \right),
\label{eq:gamma-ang}
\end{equation}
where this angle is measured in the counter-clockwise direction from the
positive $x$. We also define two special angles
\begin{equation}
\gamma_{\mathrm{b \parallel}}(\omega,\alpha) = \alpha, 
\label{eq:gamma-par} 
\end{equation}
and
\begin{equation}
\gamma_{\mathrm{b \perp}}(\omega,\alpha) = \alpha \pm 90^{\circ}.
\label{eq:gamma-perp}
\end{equation}
The first corresponds to the case when the spin velocity is directed, in the
same direction of the polarization angle $\alpha$ of the incoming beam and the
second one corresponds to the case when the spin velocity is directed
perpendicularly with respect to the polarization angle of the incoming beam.

@@

Fixing the velocity along $x$ or $y$ we define the corresponding magnitude as
\begin{align}
|\mathcal{V}^{\mathrm{a}}&(\omega,\alpha)| = \nonumber \\
&\sqrt { 
[\mathcal{V}^{\mathrm{ax}}(\omega,\alpha)]^{2} +
[\mathcal{V}^{\mathrm{ay}}(\omega,\alpha)]^{2} +
[\mathcal{V}^{\mathrm{az}}(\omega,\alpha)]^{2} .
},
\label{eq:vv-mag}
\end{align}
Then, the spin direction depends on the components of the previous equation and
so we define the spin orientation polar and azimuthal angles as
\begin{align}
\theta_{\mathrm{a}}  (\omega,\alpha)
=& 
\cos^{-1} \left( \frac{\mathcal{V}^{\mathrm{az}}(\omega,\alpha)}
{|\mathcal{V}^{\mathrm{a}}(\omega,\alpha)|} \right),
& 0 \leq &\theta \leq \pi, 
\label{eq:polar-ang}
\\
\varphi_{\mathrm{a}} (\omega,\alpha)
=& 
\tan^{-1} \left( \frac{\mathcal{V}^{\mathrm{ay}}(\omega,\alpha)}
{\mathcal{V}^{\mathrm{ax}}(\omega,\alpha)} \right),
& 0 \leq &\varphi \leq 2\pi.
\label{eq:azimuthal-ang} 
\end{align}
% where $\theta_{\mathrm{a}}(\omega,\alpha)$ is measured from the
% positive to the negative $z$ Cartesian coordinate and
% $\varphi_{\mathrm{a}}(\omega,\alpha)$ is measured on the $xy$ plane in the
% counter-clockwise direction from the positive $x$ Cartesian coordinate.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%% Theory: Layer by layer %%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Layer-by-layer analysis.}\label{sec:theory-layer}

For a layered system we have that the total contribution of Eqns. 
\eqref{eq:vs-mag} and \eqref{eq:vv-mag} are given \cite{arzatePRB14} by 
\begin{align}
|\mathcal{V}_{\sigma^{\mathrm{b}}}(\omega,\alpha)|
=& 
\sum_{\ell=1}^{N_{\mathrm{eff}}}
|\mathcal{V}_{\sigma^{\mathrm{b}}} (\ell | \omega,\alpha)|
\label{eq:vs-layer}
\\
|\mathcal{V}^{\mathrm{a}}(\omega,\alpha)|
=&
\sum_{\ell=1}^{N_{\mathrm{eff}}}
|\mathcal{V}^{\mathrm{a}} (\ell | \omega,\alpha)|
\label{eq:vv-layer}
\end{align}
where $|\mathcal{V}_{\sigma^{\mathrm{b}}} (\ell | \omega,\alpha)|$ and
$|\mathcal{V}^{\mathrm{a}} (\ell | \omega,\alpha)|$ gives the contribution of
the $\ell^{th}$ layer to the total SVI when the spin is fixed in the
$\mathrm{b}$ direction or when the velocity is fixed in the $\mathrm{a}$
direction, respectively. For the structures presented here this layers
correspond to the total number of layers being two for the \emph{up} structure
and six for the \emph{alt} structure.

@@

We preset the results for $|\mathcal{V}^{\mathrm{a}}(\omega,\alpha)|$ and
$|\mathcal{V}_{\sigma^{\mathrm{b}}}(\omega,\alpha)|$ for the
C$_{16}$H$_{8}$-alt and C$_{16}$H$_{8}$-up structures being both
noncentrosymmetric semi-infinite 2D carbon systems with 50\% hydrogenation in
different arrangements. We recall that the \emph{up} structure has hydrogen
atoms only on the upper side of the carbon sheet while the \emph{alt} structure
has alternating hydrogen atoms on the upper and bottom sides. Also we take the
hexagonal carbon lattice to be on the $xy$ plane for both structures, and the
carbon-hydrogen bonds on the perpendicular $xz$ plane, as depicted in Figs.
\ref{fig:alt-struc} and \ref{fig:up-struc}. The coordinates for the
\emph{up} and \emph{alt} unit cells of the structures are presented in Tables
\ref{tab:up-unitcell} and \ref{tab:alt-unitcell}.

>>>>

\documentclass[prb,11pt,tightenlines,twocolumn,aps]{revtex4-1}
\usepackage{amsmath}    % need for subequations 
\usepackage{graphicx}   % need for figures 
\usepackage{verbatim}   % useful for program listings 
\usepackage{color}      % use if color is used in text 
\usepackage{subfigure}  % use for side-by-side figures 
\usepackage{hyperref}   % use for hypertext links, including those to external 
                        % documents and URLs 
%\usepackage{blindtext}  % fill text 
\usepackage[dvipsnames]{xcolor}
\usepackage{multirow}
%\usepackage[capitalize]{cleveref}
% \usepackage{showframe}
\usepackage[inline]{showlabels}
\input{/Users/reinaldo/Documents/tex_files/bmsdefinitions.tex}
% \input{/Users/bms/util/definitions}

@@

\begin{abstract}
perro
%\blindtext
\end{abstract}

@@

{\color{red}
In PSCs there is no net motion of electrical charge and spin-up electrons move
in a given direction while spin-down electrons travel in the opposite one.
This effect can result from one photon absorption of linearly polarized light
by a semiconductor, with filled valence bands and empty conduction bands,
illuminated by light with photon energy larger than the energy gap.
% note 1
Using a single continuous linearly polarized laser beam, it is possible
to promote electrons in $\mathbf{k}$ space regardless the symmetry of the
system resulting in a net current equal to zero. 
% 
If the phenomena is produced in a noncentrosymmetric semiconducting media, the
electrons promoted to the conduction bands at opposite $\mathbf{k}$ points
can produce a net spin polarization different to zero\cite{alvaradoPRL85},
resulting in a PSC because the velocity of electrons
at opposite $\mathbf{k}$ points are in opposite directions.
}

%%%BMSD
The operator that describes the electronic PSC is written as
\begin{align}\label{z.1}
\hat K^{\rma\rmb}=\frac{1}{2}\left( \hat v^\rma \hat S^\rmb 
+\hat  S^\rmb \hat v^\rma\right) 
,
where $\hat\bfv=[\hat\bfr,\hat H_0]/i\hbar$ is the velocity operator, with
$\hat \bfr$ the position operator,  $\hat H_0$ the unperturbed
ground state Hamiltonian, and
the Roman superscripts  indicate Cartesian coordinates. 
To obtain the expectation value of 
$\hat K^{\rma\rmb}$, we use the length gauge for the perturbing
Hamiltonian, written as
\begin{align}\label{z.2}
\hat H_{\text{p}}=-e\hat\bfr\cdot\bfE(t)
,
where the electric field of the applied laser is given by
\begin{align}\label{z.3}
\bfE(t)= \bfE(\go)e^{-i\go t} + \bfE^*(\go)e^{i\go t}
.
In order to 
calculate the response of the system to $\bfE(t)$, one needs to
take into account the excited coherent superposition
of the spin-splited conduction bands inherent to the 
noncentrosymmetric 
semiconductors considered in this work.
%Although, this splitting is small [15,16], is required to obtain the
%correct results\cite{nastosPRB07}.
% The conduction bands in the 
% noncentrosymmetric  
% semiconductors  
% are spin split by a small amount,[15,16] typically
%   smaller than the energy width of the laser pulse, and so the pulse
%   excites a coherent superposition of the two conduction bands. Even
%   for very long pulses with narrow energy widths, dephasing effects
%   lead to an energy width of the bands large enough that spin-split
%   states can become quasidegenerate. 
To include the coherences, we follow Ref.~\onlinecite{nastosPRB05} and
use a multiple
scale approach that solves the equation of motion for the single
particle density matrix $\gr_{mn}(\bfk;t)$, leading to
\begin{align}\label{z.4}
&\frac{\partial \rho_{cc'}(\bfk)}{\partial t} =
\frac{e^{2}E^{\mathrm{a}}(\omega)E^{\mathrm{b*}}(\omega)}
{i \hbar^{2}}
\sum_{v}r^{\mathrm{a}}_{cv}(\bfk) r^{\mathrm{b}}_{vc'}(\bfk)
\left( \frac{1}{\omega - \omega_{c'v}(\bfk) - i \epsilon} 
\frac{1}{\omega - \omega_{cv}(\bfk) + i \epsilon} \right)
,
where we assumed that the conduction bands $c$ and $c'$
are quasidegenerate  states and we take $\ge\to 0$ at the end of the
calculation.  The spin-splitting of the valence ($v$) bands is very
small, and is neglected throughout this work.\cite{??} 
% 􏱈 are close to  
% one another, and that the pulse is short enough so that the energy
% width overlaps the two bands, the equation of motion can be
% solved. Leaving the details of the derivation to Ap- 
% pendix A, the result for the off-diagonal component 􏰠 􏱈, c  
%%%BMSU 
The matriz elements of any operator $\calo$ are given by
$\calo_{nm}(\bfk)=\bra{n\bfk}\hat\calo\ket{m\bfk}$,
where 
$H_{0}|n\mathbf{k}\rangle = \hbar \omega_{n}(\mathbf{k})|n\mathbf{k}\rangle$ 
with $\hbar \omega_{n}(\mathbf{k})$ the energy of the
electronic band $n$ at point $\mathbf{k}$ in the irreducible Brillouin zone
(IBZ),  $|n\mathbf{k}\rangle$ is the Bloch state, and 
$\go_{nm}(\bfk)=\go_{n}(\bfk)-\go_{m}(\bfk)$.
Using 
$\mathcal{O} = \mathrm{Tr}(\hat{\rho}\hat{\mathcal{O}})$
for the expectation value of an observable $\mathcal{O}$, 
where $\mathrm{Tr}$ denotes the trace, we obtain
\begin{align}\label{z.5}
\mathcal{O} 
%=&
%\int \frac{d^{3}k}{8\pi^{3}}\sum_{c} \langle c \mathbf{k} 
%| \hat{\rho} \hat{\mathcal{O}} |
%c\mathbf{k} \rangle \nonumber \\
\int \frac{d^{3}k}{8\pi^{3}} \sum_{cc'} \rho_{cc'}(\mathbf{k}) 
\mathcal{O}_{c'c}(\mathbf{k}),
where we used 
the closure
relationship $\sum_{n}|n\mathbf{k}\rangle \langle n\mathbf{k}| = 1$,
where $n=v,c$,
and the fact that $\rho_{vn}(\bfk)=\rho_{nv}(\bfk)=0$ for $n=c,c'$. 
Therefore, 
using  Eqs. \eqref{z.4} and \eqref{z.5},
the
rate of change of $\mathcal{O}$,
$\dot{\mathcal{O}} 
=
\mathrm{Tr} \left( \frac{\partial \hat{{\rho}}}{\partial t} \hat\calo \right)
$,
 is given by
\dot{\mathcal{O}} 
&=\frac{e^{2}}{i\hbar^{2}} \int \frac{d^{3}k}{8\pi^{3}} 
\sum'_{cc'} \mathcal{O}_{c'c}(\bfk) 
r^{\mathrm{a}}_{cv}(\bfk)  r^{\mathrm{b}}_{vc'}(\bfk)  \times
& \left( \frac{1}{\omega - \omega_{c'v}(\bfk)  - i\epsilon} - 
\frac{1}{\omega - \omega_{cv}(\bfk)  + i\epsilon} \right)
E^{\mathrm{a}}(\omega) E^{\mathrm{b*}}(\omega)
\label{eq:dotO}
.
Replacing  $\hat\calo \rightarrow \hat{K}^{\mathrm{ab}}$, in 
above expression, one can show that
\begin{equation}
\dot{K}^{\mathrm{ab}}(\omega) =
\mu^{\mathrm{abcd}}(\omega)
E^{\mathrm{c}}(\omega) E^{\mathrm{d*}}(\omega),
\label{eq:dotk}
\end{equation}
where repeated cartesian are summed, and 
\begin{equation}\label{eq:mu}
\begin{aligned}
\mu^{\mathrm{abcd}}  (\omega) &
=
\frac{\pi e^{2}}{\hbar^{2}} \int 
\frac{d^{3}k}{8 \pi^{3}} \sum'_{vcc'}
  \delta(\omega-\omega_{cv}(\bfk) 
\\
\times\mathrm{Re} & \left[ K^{\mathrm{ab}}_{cc'}(\bfk) 
r^{\mathrm{c}}_{vc'}(\bfk)   
r^{\mathrm{d}}_{cv }(\bfk)  +
(c \leftrightarrow d)  
,  
\end{aligned}
\end{equation} 
is the pseudotensor that describes the rate of change of the  PSC process in semiconductors.
To derive above we used  
$
K^{\mathrm{ab}}_{nm}(\mathbf{-k}) = K^{\mathrm{ab*}}_{nm}(\mathbf{k}). 
$
that follows from time-reversal invariance. 
The prime $'$ in the sum means that $c$ and $c'$ are quasi degenerate states and the
sum only covers these states.  Since $\mu^{\mathrm{abcd}}(\omega)$ is real we
have that $\mu^{\mathrm{abcd}}(\omega) =
\mu^{\mathrm{abdc}}(\omega)$. 
We remark that Eq.~\eqref{eq:mu} 
 is the same as Eq. (3) of Bhat et al.\cite{bhatPRL05}
obtained by using the semiconductor optical Bloch equations.
Using the closure relation,
\begin{equation}
K^{\mathrm{ab}}_{cc'}(\mathbf{k}) = \frac{1}{2}
\sum_{l=v,c}
\left(v^{\mathrm{a}}_{cl}(\mathbf{k})S^{\mathrm{b}}_{lc'}(\mathbf{k})
+S^{\mathrm{b}}_{cl}(\mathbf{k}) v^{\mathrm{a}}_{lc'}(\mathbf{k})
.
\label{eq:velspimatelem}
\end{equation}

@@

%\subsection{Spin velocity injection} % (fold)
%\label{sec:theory-pure_spin_current}
Now, we define the spin velocity injection (SVI) as
\begin{equation}\label{eq:vab-w}
\mathcal{V}^{\mathrm{ab}}(\omega) \equiv
\frac{\dot{K}^{\mathrm{ab}}(\omega)}{(\hbar/2) \dot{n}(\omega)},
\end{equation}  
that gives 
the velocity, along the direction $\mathrm{a}$, at which the spin moves polarized along the 
direction $\mathrm{b}$.
The carrier injection rate $\dot n(\go)$ is written as,\cite{nastosPRB05}
\begin{equation}
\dot{n}(\omega) =
\xi^{\mathrm{ab}}(\omega) E^{c }(\omega) E^{d*}(\omega),
\label{eq:dotn}
\end{equation}
where the tensor 
\begin{equation}\label{eq:xi}
\begin{aligned}
\xi^{\mathrm{ab}}(\omega)
&
=
\frac{2\pi e^{2}}{\hbar^{2}} \int 
\frac{d^{3}k}{8 \pi^{3}}
\\\nonumber
&\times \sum_{vc}
r^{\mathrm{a}}_{vc'}(\bfk)  
r^{\mathrm{b}}_{cv }(\bfk)  
\delta(\omega-\omega_{cv}(\bfk) 
, 
\end{aligned}
\end{equation}
is related to the imaginary part of the linear optical 
response tensor by
$\mathrm{Im}[\ge^{\rma\rmb}(\go)]=2\pi\ge_0\hbar\xi^{\rma\rmb}(\go)$.

The function $\calv^{\rma\rmb}(\go)$ allow us to quantify two very
important aspects of PSC. On one hand, we can fix the spin direction
along $\mathrm{b}$,   
and calculate the resulting electron velocity.  
On the other hand, we can fix de velocity of the electron
along $\mathrm{b}$,  and study the resulting direction along which the
spin is polarized.
To this end, the added advantage of  2D structures, besides choosing
them noncentrosymmetric, is that we can use an incoming linearly
polarized beam of light
at normal incidence, and use the  direction of the polarized  electric
field to control $\calv^{\rma\rmb}(\go)$.
Indeed, writing 
$\bfE(\omega) = E_0(\omega)(\cos\ga\,\hat\bfx+\sin\ga\,\hat\bfy)$
where $\alpha$ is the polarization angle, we obtain from
 Eq. \eqref{eq:vab-w}
that
\begin{widetext}
\begin{align}
\mathcal{V}^{\mathrm{ab}}(\omega,\alpha)
&= 
\frac{2}{\hbar\xi(\go)}
\left(\mu^{\mathrm{abxx}}(\omega)\cos^{2}\alpha + 
\mu^{\mathrm{abyy}}(\omega)\sin^{2}\alpha + 
\mu^{\mathrm{abxy}}(\omega)\sin 2\alpha\right)
\label{eq:vab-aw}
\end{align}
\end{widetext}
as for the structures chosen in this article,
$\xi^{\mathrm{xx}}(\go)=\xi^{\mathrm{yy}}(\go)\equiv\xi(\go)$, and $\xi^{\mathrm{xy}}(\go)=0$.
Now, we formalize our two options for $\calv^{\rma\rmb}$.
%

% Two interesting possibilities to analyze the SVI are fixing the spin along $z$,
% directed perpendicularly to the surface of the structure, or fixing the
% velocity along $x$ or $y$ on the $xy$ plane of the structures. Also we can
% analyze the SVI contribution coming from each layer of the structure. In
% following subsections we present these three cases.

@@

Analyzing the SVI, Eq. \eqref{eq:vab-aw}, we define the magnitude of
the electron's in plane
velocity with it's spin polarized along the $\mathrm{b}$ direction as
\begin{equation}
\mathcal{V}_{\sigma^{\mathrm{b}}}(\omega,\alpha)
\equiv
\sqrt{
[\mathcal{V}^{\mathrm{xb}}(\omega,\alpha)]^{2}\ +
[\mathcal{V}^{\mathrm{yb}}(\omega,\alpha)]^{2}\ 
}, 
\label{eq:vs-mag}
\end{equation}
and define the angle at which the velocity is directed on the $xy$ plane as
\begin{equation}
\gamma_{\gs^\mathrm{b}} (\omega,\alpha)
=
\tan^{-1} \left( \frac{\mathcal{V}^{\mathrm{yb}}(\omega,\alpha)}
{\mathcal{V}^{\mathrm{xb}}(\omega,\alpha)} \right)
.
\label{eq:gamma-ang}
\end{equation}
% where this angle is measured in the counter-clockwise direction from the
% positive $x$. 
We also define two special angles
\begin{equation}
\gamma_{\gs^\mathrm{b}}^\parallel(\omega,\alpha) = \alpha, 
\label{eq:gamma-par} 
\end{equation}
and
\begin{equation}
\gamma_{\gs\mathrm{b}}^\perp(\omega,\alpha) = \alpha \pm 90^{\circ},
\label{eq:gamma-perp}
\end{equation}
corresponding to the electron velocity being parallel or perpendicular
the incoming polarization, 
respectively. The subscript $\gs^\rmb$ denotes the spin along $\rmb$.

@@

Fixing the calculated velocity along $\rma=x$ or $\rma=y$
 we define it's corresponding magnitude as
\begin{align}
\mathcal{V}_{\mathrm{a}}&(\omega,\alpha) \equiv \nonumber \\
&\sqrt { 
[\mathcal{V}^{\mathrm{ax}}(\omega,\alpha)]^{2} +
[\mathcal{V}^{\mathrm{ay}}(\omega,\alpha)]^{2} +
[\mathcal{V}^{\mathrm{az}}(\omega,\alpha)]^{2} 
},
\label{eq:vv-mag}
\end{align}
from where we see that the spin would be oriented
in the $xyz$ system's coordinates
according to a polar angle
\begin{align}
\theta_{\mathrm{a}}  (\omega,\alpha)
=& 
\cos^{-1} \left( \frac{\mathcal{V}^{\mathrm{az}}(\omega,\alpha)}
{\mathcal{V}_{\mathrm{a}}(\omega,\alpha)} \right),
& 0 \leq &\theta \leq \pi, 
\label{eq:polar-ang}
\end{align}
and an azimuthal angle
\begin{align}
\varphi_{\mathrm{a}} (\omega,\alpha)
=& 
\tan^{-1} \left( \frac{\mathcal{V}^{\mathrm{ay}}(\omega,\alpha)}
{\mathcal{V}^{\mathrm{ax}}(\omega,\alpha)} \right),
& 0 \leq &\varphi \leq 2\pi.
\label{eq:azimuthal-ang} 
\end{align}

@@

We preset the results for 
and
$\mathcal{V}_{\sigma^{\mathrm{b}}}(\omega,\alpha)$ 
and
$\mathcal{V}_{\mathrm{a}}(\omega,\alpha)$ 
for the
C$_{16}$H$_{8}$-up  and C$_{16}$H$_{8}$-alt 
structures being both
noncentrosymmetric semi-infinite 2D carbon systems with 50\% hydrogenation in
different arrangements. We recall that the \emph{up} structure has hydrogen
atoms only on the upper side of the carbon sheet while the \emph{alt} structure
has alternating hydrogen atoms on the upper and bottom sides. Also we take the
hexagonal carbon lattice to be on the $xy$ plane for both structures, and the
carbon-hydrogen bonds on the perpendicular $xz$ plane, as depicted in Figs.
\ref{fig:up-struc} and \ref{fig:alt-struc}. The coordinates for the
\emph{up} and \emph{alt} unit cells of the structures are presented in Tables
\ref{tab:up-unitcell} and \ref{tab:alt-unitcell}.
